Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-06-15 PAPER.tex
Revised language in eigenmode convergence section

%Just as important as the eigenvalues are the eigenvectors of our empirical covariances. 
\cc{start of carina edits}The overall convergence of the covariance is important, but it is an important fact that, in general, different eigenvectors converge to their true forms at different rates. This is illustrated by Figure \ref{fig:toy_sigloss17}, which shows the convergence of eigenvectors in an empirical estimate of a covariance matrix. For this simulation, we construct a covariance whose ``true" form combines the same mock foreground from the previous toy models with an EoR component that is modeled as a diagonal matrix with eigenvalues spanning one order of magnitude (more specifically, we construct the EoR covariance as a diagonal matrix in the power spectrum domain, where the signal is expected to be uncorrelated; its Fourier-transform is then the ``true" covariance of the EoR in the visibility-domain, or $\textbf{C}_{\rm EoR}$). For different numbers of realizations, we draw random EoR signals that are consistent with $\textbf{C}_{\rm EoR}$, add them to the mock foreground data, and compute the combined empirical covariance (average over the realizations). The eigenvectors of this empirical covariance are then compared to the ``true" eigenvectors $\widehat{\textbf{v}}$, where we use as a convergence metric $\varepsilon(\widehat{\textbf{v}})$, defined as:
\begin{equation}
\label{eq:converge_eig}
\varepsilon (\widehat{\textbf{v}}) \equiv \sqrt{\sum_{i}^{N_{f}}|\textbf{v}-\widehat{\textbf{v}}|_{i}^2},
\end{equation}
where $N_{f}$ is the number of frequencies ($20$) in the mock data. The eigenmode convergence curves  in Figure \ref{fig:toy_sigloss17} are ranked ordered by eigenvalue, such that ``Eigenmode \#0" illustrates the convergence of the eigenvector with the largest eigenvalue, ``Eigenmode \#1" for the second largest eigenvalue, and so on. We see that the zeroth eigenmode is quickest to converge --- this mode describes the foreground signal. We will now explain this behavior and describe how the convergence rates of empirical eigenmodes is related to sample variance in our empirical estimate.

In general, computing an empirical covariance from a finite ensemble average means that the empirical eigenmodes have sample variances. When there are many similarly-shaped eigenmodes (i.e. modes $1$-$19$), this sample variance can lead to degeneracies between modes, where some modes are swapping with others and fluctuating about. In other words, it takes more realizations in order to sort out the degeneracies and approach the true eigenmodes. However, an eigenmode that differs substantially from the others (i.e. mode $0$, or the foreground mode), can more easily converge to its true shape because the sample variance isn't large enough for it to be mistaken for a different mode (i.e. it can decouple from the other modes more easily). Another way of thinking about this, is that for the case of a completely flat eigenspectrum (all equal eigenvalues), any linear combination of eigenvectors is allowed, and hence convergence is very difficult. The more color, or steepness, given to an eigenspectrum, the easier it is for the eigenmodes to de-couple and approach their true ones.

While the connection between the rate of convergence of an empirical eigenvector with the steepness of an eigenspectrum is interesting, it is also important to note that regardless of convergence rate, any mode that is coupled to the data is susceptible to signal loss. In the toy models throughout Section \ref{sec:SiglossOverview}, we exploit the fact that the strongest eigenmode (highest eigenvalue mode) is dominated by foregrounds in order to purposely incur signal loss for that mode. Even for the case of real PAPER data (Section \ref{sec:Sigloss}), we make the assumption that the strongest eigenmodes are likely the most contaminated by foregrounds. However, in general, foregrounds need not be restricted to the strongest eigenmodes, and as we have seen, it is really the degeneracies between eigenmodes that determines how quickly they converge, and hence how much signal loss can result.

>>>>

%Just as important as the eigenvalues are the eigenvectors of our empirical covariances. 
\cc{start of carina edits}The overall convergence of the covariance is important, but also noteworthy is the fact that different eigenvectors converge to their true forms at different rates. This is illustrated by Figure \ref{fig:toy_sigloss17}, which shows the convergence of eigenvectors in an empirical estimate of a covariance matrix. For this simulation, we construct a covariance whose true form combines the same mock foreground from the previous toy models with an EoR component that is modeled as a diagonal matrix with eigenvalues spanning one order of magnitude (more specifically, we construct the EoR covariance as a diagonal matrix in the Fourier domain, where the signal is expected to be uncorrelated; its Fourier transform is then the true covariance of the EoR in the frequency domain, or $\textbf{C}_{\rm EoR}$). For different numbers of realizations, we draw random EoR signals that are consistent with $\textbf{C}_{\rm EoR}$, add them to the mock foreground data, and compute the combined empirical covariance by averaging over the realizations. The eigenvectors of this empirical covariance are then compared to the true eigenvectors $\widehat{\textbf{v}}$, where we use as a convergence metric $\varepsilon(\widehat{\textbf{v}})$, defined as:
\begin{equation}
\label{eq:converge_eig}
\varepsilon (\widehat{\textbf{v}}) \equiv \sqrt{\sum_{i}^{N_{f}}|\textbf{v}-\widehat{\textbf{v}}|_{i}^2},
\end{equation}
where $N_{f}$ is the number of frequencies ($20$) in the mock data. The eigenmode convergence curves in Figure \ref{fig:toy_sigloss17} are ranked ordered by eigenvalue, such that ``Eigenmode \#0" illustrates the convergence of the eigenvector with the largest eigenvalue, ``Eigenmode \#1" for the second largest eigenvalue, and so on. We see that the zeroth eigenmode --- the mode describing the foreground signal --- is quickest to converge.

Our numerical test reveals that the convergence rates of empirical eigenvectors is related to the sample variance in our empirical estimate. In general, computing an empirical covariance from a finite ensemble average means that the empirical eigenmodes have sample variances. Consider first a limiting case where all eigenvalues are equal. In such a scenario, any linear combination of eigenvectors is also an eigenvector, and thus there is no sensible way to define the convergence of eigenvectors. In our current test, aside from the first mode, the eigenvalues have similar values but are not precisely equal. There are thus a well-defined set of eigenvectors to converge to. However, due to the sample variance of our empirical covariance estimate, there may be accidental degeneracies between modes, where some modes are mixing and swapping with others. Therefore, the steeper an eigenspectrum, the easier it is for the eigenmodes to de-couple from each other and approach their true forms. A particularly drastic example of this can be seen in the behavior of mode $0$ (the foreground mode), whose eigenvalue differs enough from the others that it is able to converge reasonably quickly despite substantial sample variance in our empirical covariance estimate. To break degeneracies in the remaining modes, however, requires many more realizations.

While the connection between the rate of convergence of an empirical eigenvector with the steepness of an eigenspectrum is interesting, it is also important to note that regardless of convergence rate, any mode that is coupled to the data is susceptible to signal loss. The true eigenvectors are not correlated with the data realizations; thus, if our empirical eigenvectors are converged, there will not be any signal loss. However, an unconverged eigenvector estimate will retain some memory of the data realizations used in its generation, leading to signal loss.

In the toy models throughout Section \ref{sec:SiglossOverview}, we exploit the fact that the strongest eigenmode (highest eigenvalue mode) is dominated by foregrounds in order to purposely incur signal loss for that mode. Even for the case of real PAPER data (Section \ref{sec:Sigloss}), we make the assumption that the strongest eigenmodes are likely the most contaminated by foregrounds. However, in general, foregrounds need not be restricted to the strongest eigenmodes, and as we have seen, it is really the degeneracies between eigenmodes that determines how quickly they converge, and hence how much signal loss can result.
